The Place of the Local Group in the Cosmic Web

### 2014-08-04 paper/main.tex
some edits in the abstract + introduction + new references

\begin{abstract}
In this Letter we explore the characteristics of the Local Group (LG)
location in the cosmic web using a cosmological simulation in the
$\Lambda$CDM cosmology.
We use the Hessian of the gravitational potential to define the cosmic web, 
where regions are classified as either peak, sheet, filament or void.
%We use a description for the cosmic web that uses the Hessian of the
%gravitational potential to classify a neighborhod in the simulation
%as either peak, sheet, filament or void. 
Results indicate: 
i) The LG pairs are preferentially located in filaments and sheets,
ii) There is a strong signature that LG pairs orientation depends on the cosmic web, 
in particular for pairs in filaments/sheets the orbital angular momentum tends to be perpendicular to the filament direction or the sheet plane.
iii) They are located in in a narrow range of local overdensity $0<\delta<4$, 
ellipticity $0.1<e<1.0$ and prolateness $-0.5<p<0.5$. 
%Furthermore we find a strong anti-alignment between the
%LG angular momentum vector and the filament lines and the sheet planes.  
\end{abstract}

@@

%lg rare
The configuration of galaxies in the Local Group (LG hereafter) is
quite rare to find in the local Universe and in cosmological
simulations. The LG is dominated by the two big spirals MW and M31, the next
most-luminous galaxy is M33 which is $\sim 10$ times less massive than
M31, followed by several less luminous dwarf galaxies, up to a
distance of $\sim 3$\mpc.   The velocity vector of M31, with a low
tangential velocity is consistent to a head-on collision orbit toward
the MW
\citep{2008MNRAS.386..461C,2012ApJ...753....8V,2012ApJ...753....7S}.

@@

All this combination of features make LG analogues very rare to
find. Using numerical simulatons \citet{lganalogues} found less than
$2\%$ MW-sized halos reside in a pair similar to MW-M31 and in a
similar environment. Furthermore, if we select pairs constrained
within $2\sigma$ error fron current observational measurements of the
velocity components and distance to M31, there are only $46$ systems
in a cube of $250$ \hmpc side.

\citet{2013ApJ...767L...5F} also studied MW-M31 pairs in numerical
simulations finding the typical quantities characterizing the orbital
parameters of the LG are rare among typical pairs, but not enough to
challenge the \lcdm model. 
Another definition of LG analogues is made by
\citet{2008MNRAS.384.1459L}, but despite differences ocurr in the
definitions and resulting fraction of LG analogues, all are in
agreement with a low frequency of these pairs. 
%lw08 also look lg analogues

%larger scales & lg env
To better understand the properties of the LG and how this uncommon
pair configuration fit in the cosmological context, an immediate
question arise. What else can we say of the LG at larger scales?. In
particular, which are the typical/prefered locations of these systems
within the Cosmic Web?.

Looking the LG at larger scales we have it is located in a difuse and
warped filament/wall conecting Virgo Cluster with Fornax Cluster, some
nearby galaxies and groups members of this large structure are the
Maffei group, NGC $6744$, NGC $5128$, $M101$, $M81$, $NGC1023$, Cen A
group. At this scale, there is no evident alignment of the MW-M31
orbital plane with the local filament or the Virgo-Fornax
direction. However, if we look in a smaller volume below scales of
$\sim 6$ \mpc, there is a clear alignment of th the MW-M31 orbit with
a local plane as shown by figure $3$ in \citet{2013AJ....146...69C}.

%Tweb definition
The environment definition we use in this Letter uses the tidal tensor
field.

@@

\subsection{The Bolshoi simulation}
We use the Bolshoi simulation of $\Lambda$CDM cosmology: $\Omega_{\rm
  m}=1-\Omega_{\Lambda}=0.27$, $H_0=70\,\rm km/s/Mpc$,
$\sigma_8=0.82$, $n_s=0.95$ \citep{2011ApJ...740..102K}, compatible
with the constraints from the WMAP satellite
\citep{hinshaw_etal13}. The simulation followed evolution of dark
matter in a $250 \hmpc$ box with spatial resolution of $\approx
1h^{-1}$~kpc and mass resolution of $m_{\rm p}=1.35\times 10^8\ \rm
M_{\odot}$. Halos are identified with the BDM algorithm
\citep{1997astro.ph.12217K}. The BDM algorithm is  a spherical
overdensity halo finding algorithm and is designed to identify both
host halos and subhalos.

@@

\begin{equation}
T_{ij} = \frac{\partial^2 \phi}{\partial r_i \partial r_j}, 
\end{equation}
%
where $r_{i}$, $i=1,2,3$ refers to the three spatial comoving
coordiates and $\phi$ is the gravitational potential renormalized to
follow the following Poisson equation $\nabla^2\phi=\delta$ where
$\delta$ is the matter overdensity.

@@

\cite{Tweb} performed a detailed study for the topology of the
cosmic web and its visual counterpart as a function of the parameter
$\lambda_{\rm th}$. They found that reasonable results in terms of the
volume fraction occupied by voids, the visual inspection and the halo
populations in each web type can be reached by values of $0.2<\lambda_{\rm
th}<0.4$. In this Letter we choose the value of $\lambda_{\rm
  th}=0.3$ to proceed with our analysis. We have veryfied that the
main trends reported in this paper are unsensitive to the choice of
that parameter, as long as it is in the range already quoted.

@@

After computing the eigenvalues and eigenvectors on each point of the
grid we use the matter overdensity, ellipticity and the prolatenes to
further characterize the web. These quantities are defined in terms of the
eigenvalues as follows 
%
\begin{equation}
\delta = \lambda_1 + \lambda_2 + \lambda_3,
\end{equation}
%
\begin{equation}
e= \frac{\lambda_3 - \lambda_1}{2(\lambda_1 + \lambda_2 + \lambda_3)}, 
\end{equation}
%
\begin{equation}
p= \frac{\lambda_1 + \lambda_3 - 2\lambda_2}{2(\lambda_1 + \lambda_2 +
  \lambda_3)}.
\end{equation}

@@

@@

%\begin{figure}
%\begin{center}
%  \includegraphics[width=0.48\textwidth]{alignments_e1_r_all_environments.pdf} 
%  \includegraphics[width=0.48\textwidth]{alignments_e2_r_all_environments.pdf} 
%  \includegraphics[width=0.48\textwidth]{alignments_e3_r_all_environments.pdf} 
%\end{center}
%\caption{Cumulative distributions for the alignment between the unit
%  vector linking the two halos, $\hat{r}$, and the eigenvectors
%  defining the cosmic web.
%    \label{fig:alignment_r}}  
%\end{figure}

>>>>

\begin{abstract}
In this Letter we explore the characteristics of the Local Group (LG)
location in the cosmic web using a cosmological simulation in the
$\Lambda$CDM cosmology.
We use the Hessian of the gravitational potential to define the cosmic web, 
where regions are classified as either peak, sheet, filament or
void. 
The LG in the simulations are split into two samples.
The first is a general sample composed by halo pairs with
similar masses and isolation criteria as observed in the LG. 
The second is a subset with additional kinematic constaints.
We find that the pairs in the LG sample with all constraints are:
i) Preferentially located in filaments and sheets,
ii) located in in a narrow range of local overdensity $0<\delta<4$, 
web ellipticity $0.1<e<1.0$ and prolateness $-0.5<p<0.5$. 
iii) Clearly aligned with the cosmic web, in particular for pairs in
filaments/sheets the orbital angular momentum tends to be
perpendicular to the filament direction or the sheet plane. A stronger
alignment is present for the vector linking the two halos, which lies
along the filament or the sheet plane.  
We show that the first and second results are expected trends with the
LG total mass. 
However, the strong alignment with the cosmic web cannot be explained
in the same way.

\end{abstract}

@@

%lg rare
The spatial and kinematic configuration of the Local Group
(LG) galaxies is rare to find in the local Universe and in cosmological
simulations. 
The LG is dominated by the two big spirals MW and M31, the next
most-luminous galaxy is M33 which is $\sim 10$ times less massive than
M31, followed by several less luminous dwarf galaxies, up to a
distance of $\sim 3$\mpc.   
The velocity vector of M31, with a low
tangential velocity is consistent to a head-on collision orbit toward
the MW
\citep{2008MNRAS.386..461C,2012ApJ...753....8V,2012ApJ...753....7S}.

@@

All this combination of features make LG analogues rare to
find. 
Using numerical simulatons \citet{lganalogues} found less than
$2\%$ MW-sized halos reside in a pair similar to MW-M31 and in a
similar environment. 
Furthermore, if we select pairs constrained
within $2\sigma$ error fron current observational measurements of the
velocity components and distance to M31, there are only $46$ systems
in a cubic volume of $250$ \hmpc side, giving a number density $\sim
1.0\times 10^{-6}$Mpc$^{3}$, comparable to the abundance of massive
clusters.

\citet{2013ApJ...767L...5F} also studied MW-M31 pairs in numerical
simulations finding the typical quantities characterizing the orbital
parameters of the LG are rare among typical pairs, but not enough to
challenge the \lcdm model.

Another definition of LG analogues is made by
\citet{2008MNRAS.384.1459L}, but despite differences ocurr in the
definitions and resulting fraction of LG analogues, all are in
agreement with a low frequency of these pairs. 
%lw08 also look lg analogues

%larger scales & lg env
To better understand the properties of the LG and how this uncommon
pair configuration fit in the cosmological context, an immediate
question arise. What else can we say of the LG at larger scales?. In
particular, which are the typical/prefered locations of these systems
within the Cosmic Web?. To what extent is this an expected
configuration in $\Lambda$CDM.

Looking the LG at larger scales we have it is located in a difuse and
warped filament/wall conecting Virgo Cluster with Fornax Cluster, some
nearby galaxies and groups members of this large structure are the
Maffei group, NGC $6744$, NGC $5128$, M$101$, M$81$, NGC$1023$, Cen A
group. At this scale, there is no evident alignment of the MW-M31
orbital plane with the local filament or the Virgo-Fornax
direction. However, if we look in a smaller volume below scales of
$\sim 6$ \mpc, there is a clear alignment of th the MW-M31 orbit with
a local plane as shown by Figure $3$ in \citet{2013AJ....146...69C}.

@@

\subsection{The Bolshoi simulation}
We use the Bolshoi simulation of $\Lambda$CDM cosmology: $\Omega_{\rm
  m}=1-\Omega_{\Lambda}=0.27$, $H_0=70\,\rm km/s/Mpc$,
$\sigma_8=0.82$, $n_s=0.95$ \citep{2011ApJ...740..102K}, compatible
with the constraints from the WMAP satellite
\citep{hinshaw_etal13}. The simulation followed the evolution of dark
matter in a $250 \hmpc$ box with spatial resolution of $\approx
1h^{-1}$~kpc and mass resolution of $m_{\rm p}=1.35\times 10^8\ \rm
M_{\odot}$. Halos are identified with the BDM algorithm
\citep{1997astro.ph.12217K}. The BDM algorithm is  a spherical
overdensity halo finding algorithm and is designed to identify both
host halos and subhalos.

@@

\begin{equation}
T_{ij} = \frac{\partial^2 \phi}{\partial r_i \partial r_j}, 
\end{equation}
%
where $r_{i}$, $i=1,2,3$ refers to the three spatial comoving
coordiates and $\phi$ is the gravitational potential renormalized to
follow the Poisson equation $\nabla^2\phi=\delta$ where
$\delta$ is the matter overdensity.

@@

\cite{Tweb} performed a detailed study for the topology of the
cosmic web and its visual counterpart as a function of the parameter
$\lambda_{\rm th}$. 
They found that reasonable results in terms of the
volume fraction occupied by voids, the visual inspection and the halo
populations in each web type can be reached by values of $0.2<\lambda_{\rm
th}<0.4$. 
In this Letter we choose the value of $\lambda_{\rm
  th}=0.3$ to proceed with our analysis. 
This is only relevant to the classification of the simulation into web
elements. Other results are completeley independent of this
choice. Nevertheless we have checked that the main conclusions of this
work do not depend on the choise of $\lambda_{\rm th}$.

@@

We use the matter overdensity, ellipticity and the prolatenes to
further characterize the web. 
These quantities are defined in terms of the
eigenvalues as follows 
%
\begin{equation}
\delta = \lambda_1 + \lambda_2 + \lambda_3,
\end{equation}
%
\begin{equation}
e= \frac{\lambda_3 - \lambda_1}{2(\lambda_1 + \lambda_2 + \lambda_3)}, 
\end{equation}
%
\begin{equation}
p= \frac{\lambda_1 + \lambda_3 - 2\lambda_2}{2(\lambda_1 + \lambda_2 +
  \lambda_3)}.
\end{equation}

@@

The data of the BDM halos and the Tweb is publicly available through
a database located at \url{http://www.cosmosim.org/}. A detailed
description of the database structure was presented by \cite{Riebe2013}.

@@

\begin{figure}
\begin{center}
  \includegraphics[width=0.48\textwidth]{alignments_e1_r_all_environments.pdf} 
  \includegraphics[width=0.48\textwidth]{alignments_e2_r_all_environments.pdf} 
  \includegraphics[width=0.48\textwidth]{alignments_e3_r_all_environments.pdf} 
\end{center}
\caption{Cumulative distributions for the alignment between the unit
  vector linking the two halos, $\hat{r}$, and the eigenvectors
  defining the cosmic web.
    \label{fig:alignment_r}}  
\end{figure}
